Visualize ukb whole genome, using ctwas results

### 2023-03-22 analysis/Paper_figures_simulation.Rmd
revision

```{r paramplotfunc}

plot_single <- function(mtxlist, truecol, estcol, ...){
  truth <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mean(mtxlist[[x]][, truecol]))))
  est <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mtxlist[[x]][, estcol])))

@@

  plot(est, pch = 19, xaxt = "n", xlab="" ,frame.plot=FALSE, col = colorsall[col], ...)
  axis(side=1, at=1:2, labels = c("setting 1", "setting 2"), tick = F)
  #text(x=1:length(mtxlist), 0, labels = paste0("temp",1:length(mtxlist)), xpd = T, pos =1)
  for (t in 1:nrow(truth)){
    row <- truth[t,]
    segments(row[1]-0.2, row[2] , row[1] + 0.2, row[2],
       col = colorsall[t], lty = par("lty"), lwd = 2, xpd = FALSE)
  }
  grid()
}

plot_par <- function(configtag, runtag, simutaglist){
  mtxlist <- list()
  for (group in 1:length(simutaglist)){
    simutags <- simutaglist[[group]]
    source(paste0(outputdir, "config", configtag, ".R"))
    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".s2.susieIrssres.Rd")
    susieIfs2 <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s2.susieIrss.txt")
    mtxlist[[group]] <- show_param(phenofs, susieIfs, susieIfs2, thin = thin)
  }

  # pi1 
  par(mfrow=c(1,4))
  plot_single(mtxlist, truecol = "pi1.gene_truth", estcol = "pi1.gene_est", ylab ="gene pi1", ylim = c(0,0.02), xlim = c(0.8,2.4))
  plot_single(mtxlist, truecol = "enrich_truth", estcol = "enrich_est",ylab ="gene enrichment", ylim = c(0,120), xlim = c(0.8,2.4) )
  plot_single(mtxlist, truecol = "sigma.gene_truth", estcol = "sigma.gene_est", ylab = "gene effect size", ylim = c(0.01, 0.03), xlim = c(0.8,2.4))
  plot_single(mtxlist, truecol = "PVE.gene_truth", estcol = "PVE.gene_est", ylab ="gene PVE", ylim = c(0, 0.1), xlim = c(0.8,2.4))

  return(mtxlist)
}

@@

Figure (supplementary):
```{r paramplot, fig.width= 8, fig.height= 3}
simutaglist = list(paste(4, 1:5, sep = "-"), paste(10, 1:5, sep="-"))
mtxlist <- plot_par(configtag, runtag, simutaglist)
```

Back up figures

I have in total 10 x 2 = 20 settings to use generate the parameter figures.

### Comparison of PVE.gene vs. MESC estimates.
Leave it to Kevin.

@@

@@

@@

  layout(matrix(1:2, ncol = 1), widths = 1, heights = c(1.5,1.5), respect = FALSE)
  par(mar = c(0, 4.1, 4.1, 2.1))
  plot(a[a$type=="SNP"]$p0, a[a$type == "SNP"]$PVALUE, pch = 19, xlab="Genomic position" ,frame.plot=FALSE, col = "white", ylim= c(-0.1,1.1), ylab = "ctwas PIP", xaxt = 'n')
  grid()
  points(a[a$type=="SNP"]$p0, a[a$type == "SNP"]$susie_pip, pch = 21, xlab="Genomic position", bg = colorsall[1])
  points(a[a$type=="SNP" & a$r2max > r2cut, ]$p0, a[a$type == "SNP"  & a$r2max >r2cut]$susie_pip, pch = 21, bg = "purple")
  points(a[a$type=="SNP" & a$ifcausal == 1, ]$p0, a[a$type == "SNP" & a$ifcausal == 1]$susie_pip, pch = 21, bg = "salmon")
  points(a[a$type=="gene" ]$p0, a[a$type == "gene" ]$susie_pip, pch = 22, bg = colorsall[1], cex = 2)
  points(a[a$type=="gene" & a$r2max > r2cut, ]$p0, a[a$type == "gene"  & a$r2max > r2cut]$susie_pip, pch = 22, bg = "purple", cex = 2)
  points(a[a$type=="gene" & a$ifcausal == 1, ]$p0, a[a$type == "gene" & a$ifcausal == 1]$susie_pip, pch = 22, bg = "salmon", cex = 2)
  if (isTRUE(plot_ceqtl)){
      for (cgene in a[a$type=="gene" & a$ifcausal == 1, ]$id){
     load(paste0(tools::file_path_sans_ext(exprfs[a[a$id == cgene, "chrom.x"][[1]]]), "qc.Rd"))
     eqtls <- rownames(wgtlist[[cgene]])
     points(a[a$id %in% eqtls,]$p0, rep( -0.15, nrow(a[a$id %in% eqtls,])), pch = "|", col = "salmon", cex = 1.5)
    }
  }

  if (isTRUE(plot_eqtl)){
    for (cgene in a[a$type=="gene" & a$PVALUE >5, ]$id){
     load(paste0(tools::file_path_sans_ext(exprfs[a[a$id == cgene, "chrom.x"][[1]]]), "qc.Rd"))
     eqtls <- rownames(wgtlist[[cgene]])
     points(a[a$id %in% eqtls,]$p0, rep( -0.15, nrow(a[a$id %in% eqtls,])), pch = "|", col = "black", cex = 1.5)
    }
  }

@@

>>>>

```{r paramplotfunc}
require(latex2exp)
plot_single <- function(mtxlist, truecol, estcol, xlabels = c("setting 1", "setting 2"), ...){
  truth <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mean(mtxlist[[x]][, truecol]))))
  est <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mtxlist[[x]][, estcol])))

@@

  plot(est, pch = 19, xaxt = "n", xlab="" ,frame.plot=FALSE, col = colorsall[col], ...)
  axis(side=1, at=1:2, labels = xlabels, tick = F)
  #text(x=1:length(mtxlist), 0, labels = paste0("temp",1:length(mtxlist)), xpd = T, pos =1)
  for (t in 1:nrow(truth)){
    row <- truth[t,]
    segments(row[1]-0.2, row[2] , row[1] + 0.2, row[2],
       col = colorsall[t], lty = par("lty"), lwd = 2, xpd = FALSE)
  }
  grid()
}

get_params <- function(configtag, runtag, simutaglist){
    mtxlist <- list()
  for (group in 1:length(simutaglist)){
    simutags <- simutaglist[[group]]
    source(paste0(outputdir, "config", configtag, ".R"))
    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".s2.susieIrssres.Rd")
    susieIfs2 <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s2.susieIrss.txt")
    mtxlist[[group]] <- show_param(phenofs, susieIfs, susieIfs2, thin = thin)
  }
    return(mtxlist)
}

plot_par <- function(mtxlist){
  par(mfrow=c(1,4))
  plot_single(mtxlist, truecol = "pi1.gene_truth", estcol = "pi1.gene_est", ylab ="gene pi1", ylim = c(0,0.02), xlim = c(0.8,2.4))
  plot_single(mtxlist, truecol = "enrich_truth", estcol = "enrich_est",ylab ="gene enrichment", ylim = c(0,120), xlim = c(0.8,2.4) )
  plot_single(mtxlist, truecol = "sigma.gene_truth", estcol = "sigma.gene_est", ylab = "gene effect size", ylim = c(0.01, 0.03), xlim = c(0.8,2.4))
  plot_single(mtxlist, truecol = "PVE.gene_truth", estcol = "PVE.gene_est", ylab ="gene PVE", ylim = c(0, 0.1), xlim = c(0.8,2.4))
}

@@

## Parameters Figure
```{r paramplot, fig.width= 8, fig.height= 3}
simutaglist = list(paste(4, 1:5, sep = "-"), paste(10, 1:5, sep="-"))
mtxlist <- get_params(configtag, runtag, simutaglist)
plot_par(mtxlist)
```

### Supplementary-parameters
```{r paramplot-suppl, fig.width= 8, fig.height= 3}
plot_par_snps <- function(mtxlist){
  par(mfrow=c(1,3))
  plot_single(mtxlist, truecol = "pi1.SNP_truth", estcol = "pi1.SNP_est", xlabels = c("high gene PVE", "low gene PVE"), ylab =TeX('Percent causal, $\\pi_V$'), ylim = c(0,0.0005), xlim = c(0.8,2.4))
  plot_single(mtxlist, truecol = "sigma.SNP_truth", estcol = "sigma.SNP_est",xlabels = c("high gene PVE", "low gene PVE"), ylab = "Variant effect size", ylim = c(0.01, 0.03), xlim = c(0.8,2.4))
  plot_single(mtxlist, truecol = "PVE.SNP_truth", estcol = "PVE.SNP_est",xlabels = c("high gene PVE", "low gene PVE"), ylab ="Variant PVE", ylim = c(0, 0.8), xlim = c(0.8,2.4))
}

plot_par_snps(mtxlist)
```

@@

### Supplementary PIP plot

```{r pipplot, fig.width= 8, fig.height= 16}
simutaglist = lapply(c(1:3,5:9), function(x) paste(x, 1:5, sep ="-"))
mtxlist <- get_params(configtag, runtag, simutaglist)
simutaglist = c(list(paste(1, c(1,2,5), sep = "-")), 
                lapply(c(2,3,5,6,7,8,9), function(x) paste(x, 1:5, sep ="-")))

pdf(file = "temp3.pdf", height = 10, width =10)
plotlist <- list()
for (i in 1:length(simutaglist)){
  plotlist[[i]] <- plot_PIP(configtag, runtag, simutaglist[[i]], main = 
                              TeX(sprintf("\\overset{$\\pi_G$= %s, $PVE_G$ = %s;}{$\\pi_V$= %s, $PVE_G$ = %s}",
                                  signif(mean(mtxlist[[i]][,"pi1.gene_truth"]), digits=2), 
                                  signif(mean(mtxlist[[i]][,"PVE.gene_truth"]), digits=2),
                                  signif(mean(mtxlist[[i]][,"pi1.SNP_truth"]), digits=2),
                                  signif(mean(mtxlist[[i]][,"PVE.SNP_truth"]), digits=2))))
}
gridExtra::grid.arrange(plotlist[[1]], plotlist[[2]],
                        plotlist[[3]], plotlist[[4]],
                        plotlist[[5]], plotlist[[6]],
                        plotlist[[7]], plotlist[[8]],ncol = 3)
dev.off()
```

@@

## Scatter plot
```{r scatterplot}
runtag <- "ukb-s80.45-adi"
simutags <- paste(3, 2, sep = "-")
phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
gwasfs <- paste0(outputdir, runtag, "_simu",simutags, ".exprgwas.txt.gz")
pipfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
df <- scatter_plot_PIP_p(phenofs, pipfs, gwasfs, main ="PIP-p")
```

@@

  layout(matrix(1:2, ncol = 1), widths = 1, heights = c(1.5,1.5), respect = FALSE)
  par(mar = c(0, 4.1, 4.1, 2.1))
  plot(a[a$type=="SNP"]$p0, a[a$type == "SNP"]$PVALUE, pch = 19, xlab="Genomic position" ,frame.plot=FALSE, col = "white", ylim= c(-0.1,1.1), ylab = "ctwas PIP", xaxt = 'n')
  grid()
  points(a[a$type=="SNP"]$p0, a[a$type == "SNP"]$susie_pip, pch = 21, xlab="Genomic position", bg = colorsall[1])
  points(a[a$type=="SNP" & a$r2max > r2cut, ]$p0, a[a$type == "SNP"  & a$r2max >r2cut]$susie_pip, pch = 21, bg = "purple")
  points(a[a$type=="SNP" & a$ifcausal == 1, ]$p0, a[a$type == "SNP" & a$ifcausal == 1]$susie_pip, pch = 21, bg = "salmon")
  points(a[a$type=="gene" ]$p0, a[a$type == "gene" ]$susie_pip, pch = 22, bg = colorsall[1], cex = 2)
  points(a[a$type=="gene" & a$r2max > r2cut, ]$p0, a[a$type == "gene"  & a$r2max > r2cut]$susie_pip, pch = 22, bg = "purple", cex = 2)
  points(a[a$type=="gene" & a$ifcausal == 1, ]$p0, a[a$type == "gene" & a$ifcausal == 1]$susie_pip, pch = 22, bg = "salmon", cex = 2)
  if (isTRUE(plot_ceqtl)){
    gi <- 0
      for (cgene in a[a$type=="gene" & a$ifcausal == 1, ]$id){
     load(paste0(tools::file_path_sans_ext(exprfs[a[a$id == cgene, "chrom.x"][[1]]]), "qc.Rd"))
     eqtls <- rownames(wgtlist[[cgene]])
     points(a[a$id %in% eqtls,]$p0, rep( -0.15 - gi, nrow(a[a$id %in% eqtls,])), pch = "|", col = "salmon", cex = 1.5)
     gi <- gi + 0.2
    }
  }

  if (isTRUE(plot_eqtl)){
     gi <- 0
    for (cgene in a[a$type=="gene" & a$PVALUE >4, ]$id){
     load(paste0(tools::file_path_sans_ext(exprfs[a[a$id == cgene, "chrom.x"][[1]]]), "qc.Rd"))
     eqtls <- rownames(wgtlist[[cgene]])
     points(a[a$id %in% eqtls,]$p0, rep( -0.15-gi, nrow(a[a$id %in% eqtls,])), pch = "|", col = "black", cex = 1.5)
     gi <- gi + 0.2
    }
  }

@@

### coloc false positives
```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 64000000 , endpos =66000000, rerun_ctwas = F, plot_eqtl = T )
```
```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 66000000 , endpos =70000000, rerun_ctwas = F, plot_eqtl = T )
```
```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 155000000 , endpos =170000000, rerun_ctwas = F, plot_eqtl = T )
```
```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 5, startpos = 47000000 , endpos =50000000, rerun_ctwas = F, plot_eqtl = T )
```

### 2023-03-22 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
revision

plot_all <- function(susief, pf, phenof, chr, ldres = NULL, thin = 1){
  a <- fread(susief, header =T)
  setnames(a, old = "id", new = "name")
  b1 <- fread(paste0(pf, ".snpgwas.txt.gz"), header =T)
  b2 <- fread(paste0(pf, ".exprgwas.txt.gz"), header =T)
  setnames(b1, old = "pos", new = "p0")
  b <- rbind(b1, b2, fill = T)
  setnames(b, old = c('PVALUE', 'id'),
         new = c('pvalue', 'name'))

@@

@@

  fig <- subplot(fig1, fig3, fig4, fig6, nrows=4, shareX = T, titleY = T)

@@

>>>>

plot_all <- function(susief, pf, phenof, colocf = NULL, chr, ldres = NULL, thin = 1){
  a <- fread(susief, header =T)
  setnames(a, old = "id", new = "name")
  b1 <- fread(paste0(pf, ".snpgwas.txt.gz"), header =T)
  b2 <- fread(paste0(pf, ".exprgwas.txt.gz"), header =T)
  setnames(b1, old = "pos", new = "p0")
  b <- rbind(b1, b2, fill = T)
  setnames(b, old = c('PVALUE', 'id'),
         new = c('pvalue', 'name'))

@@

  if (!is.null(colocf)) {
    a <- fread(colocf, header = T)
    setnames(a, old = "ID", new = "name")
    dt.gene <- merge(dt.gene, a, by ="name")
  }

@@

  if (is.null(colocf)) {
     fig <- subplot(fig1, fig3, fig4, fig6, nrows=4, shareX = T, titleY = T)
  } else {
     fig7 <- plot_coloc_pp4(dt.gene)
     fig <- subplot(fig1, fig3, fig4, fig6, fig7, nrows=5, shareX = T, titleY = T)

  }

@@

# region plot with coloc results
```{r sim 2, out.width = "98%", out.height="95%"}
runtag <- "ukb-s80.45-adi"
i <- 4
chrom <- 5
comparedir <- "/project2/mstephens/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/"
colocfs <- paste0(comparedir, runtag, "_simu", simutags,".Adipose_Subcutaneous.coloc.result")
plot_all(susiefs[i], pfs[i], phenofs[i], chrom, colocf = colocfs[i], thin = 0.01)
```
